Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-10-03 paper/article.tex
Paper progress

\begin{abstract}
\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
intelligence (AI) and closed-loop neurostimulation to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning stimulation policies that optimize for
(a) specific regimes of neural activities, or (b) external task performance.
For example, it may seek to learn a complex stimulation policy, to aid a
stroke victim in reaching to grasp an object.
Through the use of artificial neural networks (ANNs) and deep learning, the
co-processor co-adapts with the neural circuit, allowing it to seek optimal
stimulation policies, and adapt them as the neural circuit changes. The results
presented here demonstrate a neural co-processor for the first time, through
the use of a simulation, and explore some of the core algorithms that may allow them
to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
co-processor, through the use of a simulated neural circuit, which  performs a
reach-to-grasp task, based on visual input. It is designed to closely resemble
a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate a
variety of lesions by altering the model, and
then demonstrate a co-processor's ability to restore lost function through
``stimulation'' of that model. We further test the ability of a
co-processor to adapt its stimulation as the simulated brain
undergoes changes. \textit{Main results} Our simulated co-processor
successfully co-adapts with the neural circuit to accomplish the external reaching
task. The co-processor framework demonstrated here adapts to a variety of lesion
types, and to ongoing changes in the simulated brain. \textit{Significance} The proof-of-concept here
outlines a co-processor model, as well as our approach to training it, leading to
insights on how such a model may be developed for \textit{in vivo} use. We believe this
co-processor design will allow for learning complex stimulation policies that
help restore function to a stroke victim. This initial proof-of-concept lays
the foundation for a future demonstration \textit{in vivo}.
\end{abstract}

@@

\section{Introduction}
* Aided in part by application of advanced AI techniques, BCIs have made
advancements over the last several decads, allowing for decoded
brain signals to be used for control of a wide variety of virtual and physical
prostheses.
** Separately: advances in stimulation techniques and modeling have allowed us to probe neural circuit dynamics and learn to better drive neural circuits towards target dynamics (citations)
** Increasing interesting in combining decoding and encoding, for closed-loop control. BBCIs
** They have the potential to apply real-time, fine-grained control of neural circuits.
** Allows stimulation to be conditioned on a user's intention, sensory feedback, error signals, etc., to drive e.g. improvement in motor control after impairment due to TBI.

* Introduce the notion of a co-processor
** Use of AI and deep learning to determine a stimulation policy.
** Co-adapts policy to solve a task, together with the neural circuit
** Flexible model, allows for optimization of either brain activities or external task
** "These ``neural co-processors'' can be used to jointly optimize cost functions with the
   nervous system to achieve desired behaviors ranging from targeted neuro-rehabilitation
   to augmentation of brain function."
** Proposed approach here combines a learned stimulation model with a stimulation agent
** Learning allows for active adaptation to the neural circuit as it changes.

* BCIs: a single direction
* Increasingly sophisticated ML techniques applied to decoding
** See 2019Raj citations
* Encoding models, stimulation models (cite)

* Open loop control examples.  \cite{khanna.openloop}

* Learning stimulation models \cite{shanechi.stimmodel}

* Bi-directional BCIs (BBCIs)
** Closed loop control
\begin{itemize}
	\item \cite{walker.inception} Inception Loops: driving brain states
	\item \cite{kahana.biomarker} “Closed loop” here refers to when to apply stimulation, always of the same type and at the same site, based on memory performance.
	\item \cite{tafazoli.acls} ACLS
\end{itemize}

@@

* Overview of test bed
** Michaels model \cite{michaels.mrnn}
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing

@@

@@

* White noise and noised params
** Learning instability
** Predictive power isn't enough

>>>>

\begin{abstract}
\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
intelligence (AI) for closed-loop neurostimulation to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning stimulation policies that optimize for
(a) specific regimes of neural activity, or (b) external task performance.
For example, it may seek to learn a complex stimulation policy, to aid a
stroke victim in reaching to grasp an object. Through the use of artificial neural
networks (ANNs) and deep learning, the co-processor co-adapts with the
neural circuit, allowing it to seek optimal stimulation policies, and adapt them
as the neural circuit changes. The results presented here demonstrate a
neural co-processor for the first time, through the use of a simulation, and
explore some of the core algorithms that may allow them to succeed.
\textit{Approach} We provide the first proof-of-concept of a neural co-processor,
through the use of a simulated neural circuit, which  performs a reach-to-grasp task,
based on visual input. It is designed to closely resemble a similar circuit in
a primate brain \cite{michaels.mrnn}. We simulate a variety of lesions by altering
the model, and then demonstrate a co-processor's ability to restore lost function
through ``stimulation'' of that model. We further test the ability of a co-processor to
adapt its stimulation as the simulated brain undergoes changes.
\textit{Main results} Our simulated co-processor successfully co-adapts with the neural
circuit to accomplish the external reaching task. The co-processor framework
demonstrated here adapts to a variety of lesion types, and to ongoing changes in the
simulated brain.
\textit{Significance} The proof-of-concept here outlines a co-processor model, as well
as our approach to training it, leading to insights on how such a model may be
developed for \textit{in vivo} use. We believe this co-processor design will allow for
learning complex stimulation policies that help restore function to a stroke victim.
\end{abstract}

@@

\section{Introduction}
Aided in part by application of advanced AI techniques, brain-computer interfaces (BCIs) have made
advancements over the last several decads, allowing for decoded brain signals to be used for
control of a wide variety of virtual and physical prostheses \cite{rao.bcibook} \cite{wolpaw.bcibook} \cite{moritz.neuro}
\cite{lebedev.bmi}. Separately: advances in stimulation techniques and modeling have allowed us to probe
neural circuit dynamics \textbf{TODO: citations} and learn to better drive neural circuits towards target
dynamics, by encoding and delivering information through stimulation \textbf{TODO: citations}.
Recently, there has been increasing interest in building on these advances to combine decoding
and encoding in a single system, for closed-loop stimulation of a neural circuit. Bi-directional
BCIs (BBCIs) allow stimulation to be conditioned by decoded brain activity as well as external
sensor data (e.g. camera), which can allow for the application of real-time, fine-grained control of
neural circuits and prosthetic devices \textbf{TODO: citations}. These may lead, for example, to neuro-
prostheses that are capable of restoring movement which was lost due to traumatic brain injury (TBI),
to a degree not previously possible.

Motivated by that progress, we demonstrate here a flexible and unifying framework for combining encoding
and decoding, which we term ``neural co-processors'' \cite{rao.coproc}. Neural co-processors leverage
AI and deep learning to identify optimal, closed-loop stimulation patterns. The approach is flexible
enough to optimize not only for particular neural activities, but also for tasks external to the
subject. For example, they may be able to aid a stroke victim by finding a stimulation pattern of
the motor cortex which helps restore lost limb function.

Additionally, the co-processor framework allows a neuro-prosthesis to actively adapt to
a neural circuit as it changes with time. This framework is capable of co-adapting with
the circuit, i.e. brain, by updating its stimulation regime, while at the same time the
brain is updating its response to the stimulation, and changing due to natural plasticity,
aging, etc. This allows the co-processor to continually optimize for the intended cost
function, despite the signficant non-stationarity of the target circuit.

Here we provide a proof-of-concept in simulation for a co-processor that restores
movement to a limb, after a subject has suffered a stroke affecting its ability to
use that limb. It combines:
\begin{itemize}
	\item A stimulation model, which models the relationship between decoded brain activity,
	      stimulation, and task performance.
	\item An AI agent which determines the stimulation to apply, in real time.
\end{itemize}
Proposed approach here combines a learned stimulation model with a stimulation agent

* Encoding models \cite{shanechi.stimmodel} (others?)
** Decades of advancement in encoding models, used for:
*** Better understanding the brain
*** Learning to understand the effects fo stimulation regimes
** Optogenetic vs TMS vs eCog
** In our case, designed to aid in the training of the stimulation agent.

* Open loop control
** Example: \cite{khanna.openloop}

* Bi-directional BCIs (BBCIs)
** Closed loop control
\begin{itemize}
	\item \cite{walker.inception} Inception Loops: driving brain states
	\item \cite{kahana.biomarker} “Closed loop” here refers to when to apply stimulation, always of the same type and at the same site, based on memory performance.
	\item \cite{tafazoli.acls} ACLS
	\item cite Bolus
\end{itemize}

@@

* Overview of test bed
** Michaels model \cite{michaels.mrnn}
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing
** Observation model

@@

* Training
** Alternate training EN and CPN
*** Train an EN
**** Training the EN until predictive power reaches a threshold, based on current task performance
**** Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
**** Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
**** Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
*** Train the CPN
**** Backprop through the EN, effectively using the EN's prediction as ground truth

@@

* Split by lesion design

* Show example where AIP/F5 lesioned and how that affects performance

* Any point in showing sensitivity to observation or stim dim?

* Try regularization variation?

* White noise and noised params
** Learning instability
** Predictive power isn't enough
** Illustrate this

* Training efficiency analysis
